---
title: "MLS Salaries"
author: "Liam Gesner"
date: "2022-08-16"
output: html_document
---
Loading in data

```{r}
library(tidyverse)
library(scales)
library(readr)
library(fs)
library(dplyr)
library(ggplot2)
```

Reading file paths
```{r}
file_paths <- fs::dir_ls("C:/Users/Liam/Documents/GitHub/Projects/Projects/MLS")
file_paths
```

For loop to read in file contents as a list of data frames.
```{r}
file_contents <- list()

for(i in seq_along(file_paths)){
  file_contents[[i]] <- read.csv(
    file = file_paths[[i]]
  )
}
```

Add year column to each file to distinguish between datasets.
```{r}
for(i in seq_along(file_contents)){
  Year <- 2006
  Year <- Year +i
  file_contents[[i]]$Year <- Year
}
```

Join the list of datasets into a single data set
```{r}
MLS_Salaries <- bind_rows(file_contents)
```

Grouping for analysis of MLS over recent years.
```{r}
Min_Max_Salary_By_Year <- MLS_Salaries %>%
  group_by(`Year`) %>%
  summarise(
    minimum_Salary = paste0("$",formatC(min(`guaranteed_compensation`, na.rm = T), format = "f", digits = 2, big.mark = ",")),
    maximum_Salary = paste0("$",formatC(max(`guaranteed_compensation`, na.rm = T), format = "f", digits = 2, big.mark = ","))
  )
head(Min_Max_Salary_By_Year)
tail(Min_Max_Salary_By_Year)

Clubs_by_Year <- MLS_Salaries %>%
  group_by(`Year`) %>%
  summarise(
    number_of_clubs = length(unique(`club`))
  )
head(Clubs_by_Year)
tail(Clubs_by_Year)

Salary_By_Year <- MLS_Salaries %>%
  group_by(`Year`) %>%
  summarise(
    mean_Salary = paste0("$",formatC(mean(`guaranteed_compensation`, na.rm = T), format = "f", digits = 2, big.mark = ",")),
    total_Salary = paste0("$",formatC(sum(`guaranteed_compensation`, na.rm = T), format = "f", digits = 2, big.mark = ",")),
    number_of_players = n()
  )
head(Salary_By_Year)
tail(Salary_By_Year)

C_Salary_By_Year <- MLS_Salaries %>%
  group_by(`Year`, `club`) %>%
  summarise(
    number_of_players = n(),
    median_Salary = paste0("$",formatC(median(`guaranteed_compensation`, na.rm = T), format = "f", digits = 2, big.mark = ","))
  ) 
C_Salary_By_Year <- C_Salary_By_Year %>% filter(number_of_players > 9)
head(C_Salary_By_Year)
tail(C_Salary_By_Year)

P_Salary_By_Year <- MLS_Salaries %>%
  group_by(`Year`, `position`) %>%
  summarise(
    number_of_players = n(),
    median_Salary = paste0("$",formatC(median(`guaranteed_compensation`, na.rm = T), format = "f", digits = 2, big.mark = ",")),
    mean_Salary = paste0("$",formatC(mean(`guaranteed_compensation`, na.rm = T), format = "f", digits = 2, big.mark = ","))
  ) 
P_Salary_By_Year <- P_Salary_By_Year %>% filter(number_of_players > 40)
head(P_Salary_By_Year)
tail(P_Salary_By_Year)
```

Plot club growth in MLS over the last 10 years from 2017 showing
league's ability to expand at a greater rate than other 
clubs folding.
```{r club_growth, echo=FALSE}
p_club_growth <- ggplot(data = Clubs_by_Year, aes(x=as.character(`Year`), y=`number_of_clubs`, group = 1))+
  geom_bar(stat = "identity", color = "blue", fill = "white")+
  labs(x= "Year", y = "Clubs in MLS")+
  geom_text(aes(label = `number_of_clubs`)) +
  theme_classic()
plot(p_club_growth, type = "l")
```

Despite growth of the league, the league minimum salary has more than quadrupled from 2007 to 2017 showing sustainability. 

```{r min_max, echo=FALSE}
p_min_max <- ggplot(data = Min_Max_Salary_By_Year, aes(x=as.character(`Year`), y=`minimum_Salary`, group = 1))+
  geom_bar(stat = "identity", fill = "black", color="lightgreen")+
  labs(x= "Year", y = "Minimum Salary")+
  theme_classic()
plot(p_min_max, type = "l")
```

As MLS has expanded in size as well as benefits for players, the middle 50% of player's
salaries has also at least doubled for each position. This is imperative in 
competing for better players within our region, North and South America, to improve the quality 
of the league and players.
```{r position_salary, echo=FALSE}
p_position_salary <- ggplot(data = P_Salary_By_Year, aes(x=as.character(`Year`), y=parse_number(`median_Salary`),color = `position`, group = `position`))+
  geom_line(size = 2)+
  labs(x= "Year", y = "Median Salary")+
  theme_classic()
plot(p_position_salary, type = "l")
```